\label{sec:5}
This thesis proposes a method for elastic shape mean estimation of open planar
curves. 
It combines the concept of the full Procrustes mean with the SRV framework to
estimate elastic full Procrustes means.  
This mean can be shown to be related to an eigenfunction problem over the
complex covariance surface of the observed curves, which can be estimated using
established methods for Hermitian covariance smoothing.  
These methods, in combination with the methods for elastic mean estimation in
\cite{Steyer2021}, make the proposed estimation procedure especially suitable to
sparse and irregular observations.
The method and its implementation were validated over toy datasets, where it was
shown that the estimated shape means are elastic and invariant with respect to
shape preserving transformations of the input curves.
The penalty used in the estimation of the complex covariance
surface provides stable estimates in regions where observations are sparse and
consistently  good results were achieved for first and second order penalties.
The estimated mean may be plotted together with the elastic full Procrustes
fits, however, care has to be taken when interpreting these plots, as the
estimated mean curve is not a simple functional mean of the aligned curves.
Furthermore, the elastic full Procrustes mean is quite robust to
outliers, as they tend to get shrunk towards zero, restricting their influence
on the mean estimation.
Finally, the method was used to analyse variability in an empirical dataset of
tongue contours during consonant articualtion.
Here it was found that the vowel context has the largest influence on the tongues
shape, while the effect of the spoken consonant seems to vary between speakers.
The proposed methods might be improved by including some form of smoothing into
the estimation of the elastic full Procrustes fits, such as a variation of the
curve smoothing proposed in \cref{app:a-smooth}, or by directly solving the
warping optimisation over the analytical solution to the rotation and scaling
alignment as proposed in \cref{app:a-warp}.
Two important extension would be elastic full Procrustes mean estimation for
closed curves and for higher-dimensional curves.
